% FGCS
\documentclass[final,5p,times,twocolumn]{elsarticle}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{hyperref}
\usepackage{soul}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{makeidx} 
\usepackage[section]{placeins}
\usepackage{float}


% Commands
\newcommand{\note}[2][inline]{\todo[color=yellow!30,#1]{\small\sf #2}}
\newcommand{\comment}[2][inline]{\todo[color=red!30,#1]{\small\sf #2}}

\journal{Future Generation Computer Systems}

\begin{document}
\begin{frontmatter}

% Title
\title{A Semantic-Based Approach to Attain Reproducibility of Computational Environments in Scientific Workflows: A Case Study}

% Authors and Institutions
\author[upm]{Idafen Santana-Perez\corref{cor1}}
\ead{isantana@fi.upm.es}

\author[isi]{Rafael Ferreira da Silva}
\ead{rafsilva@isi.edu}

\author[isi]{Mats Rynge}
\ead{rynge@isi.edu}

\author[isi]{Ewa Deelman}
\ead{deelman@isi.edu}

\author[upm]{Mar\'ia~S.~P\'erez-Hern\'andez}
\ead{mperez@fi.upm.es}

\author[upm]{Oscar Corcho}
\ead{ocorcho@fi.upm.es}

\cortext[cor1]{Corresponding address: Ontology Engineering Group (OEG), Laboratorio de Inteligencia Artificial, Facultad de Inform\'atica, Universidad Polit\'ecnica de Madrid, Avda. Montepr\'incipe, s/n, Boadilla del Monte, 28660, Spain}

\address[upm]{Ontology Engineering Group, Universidad Polit\'ecnica de Madrid, Madrid, Spain }
\address[isi]{University of Southern California, Information Sciences Institute, Marina del Rey, CA, USA}


% Abstract
\begin{abstract}
Reproducible research in scientific workflows is often addressed by tracking the provenance of the produced results. While this approach allows inspecting intermediate and final results, improves understanding, and permits replaying a workflow execution, it does not ensure that the computational environment is available for subsequent executions to reproduce the experiment. In this work, we propose describing the resources involved in the execution of an experiment using a set of semantic vocabularies, so as to conserve the computational environment. We define a process for documenting the workflow application, management system, and their dependencies based on 4 domain ontologies. We then conduct an experimental evaluation using a real workflow application on an academic and a public Cloud platform. Results show that our approach can reproduce an equivalent execution environment of a predefined virtual machine image on both computing platforms.
\end{abstract}

\begin{keyword}
Semantic Metadata \sep Scientific Workflow \sep Reproducibility
\end{keyword}


\end{frontmatter}


% Introduction
\section{Introduction}
\input{sec-introduction}


% Semantic Modeling
\section{Semantic Modeling of Computational Resources}
\label{sec:semantic}
\input{sec-semantic}


% Section
\section{Reproducibility in Scientific Workflows}
\label{sec:reproducibility}
\input{sec-reproducibility}


% Section
\section{Experimental Evaluation}
\label{sec:experiment}
\input{sec-experiment}


% Section
\section{Related Work}
\label{sec:related-work}
\input{sec-related-work}


% Section
\section{Conclusion and Future Work}
\label{sec:conclusion}
\input{sec-conclusion}


% Acknowledgements
\section*{Acknowledgements}

This material is based upon work supported in part by the National Science Foundation under Grant No. 0910812 to Indiana University for ``FutureGrid: An Experimental, High-Performance Grid Test-bed'', the FPU grant from the Spanish Science and Innovation Ministry (MICINN), and the Ministerio de Econom\'ia y Competitividad (Spain) project "4V: Volumen, Velocidad, Variedad y Validez en la Gesti\'on Innovadora de Datos" (TIN2013-46238-C4-2-R). We also thank Gideon Juve and Karan Vahi for their valuable help.


% Bibliography
\bibliographystyle{elsarticle-num}
\bibliography{references}

\end{document}